Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra

### 2025-05-18 paper/paper_draft.tex
paper changes

% !TEX root = /Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/paper_draft.tex
\documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers

% Packages to enhance functionality
\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{authblk}          % For author affiliations
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{subcaption}       % For subfigures within a figure
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{authblk}
\usepackage{multirow} % needed for merging cells vertically
\usepackage{booktabs}

@@

\usepackage[backend=biber, style=numeric]{biblatex}
\addbibresource{/Users/jonasfrugte/Desktop/Research_Project/fisher_calc_weak_lensing/paper/sources.bib}

@@

% Additional customizations (optional)
\setlength{\parindent}{0pt}   % No paragraph indentation
\setlength{\parskip}{1em}     % Add space between paragraphs

\title{\huge \textbf{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}}

\author[ ]{\textbf{Jonas Frugte}}
\author[1]{\textbf{P. Daniel Meerburg}}

% \affil[1]{Independent Researcher, The Netherlands}
\affil[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
\affil[ ]{}
\affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}
 % Removes the default date

\begin{document}
\maketitle

@@

\begin{center}
    \textbf{NB: Currently still in draft.}
\end{center}
\hrule

\begin{abstract}
    % Weak gravitational lensing of the Cosmic Microwave Background (CMB) and distant galaxies provides a powerful probe of the large-scale structure and cosmic expansion history. Upcoming Stage-IV surveys like the Simons Observatory, CMB-S4, LSST, and Euclid promise unprecedented precision, motivating detailed forecasts of their constraining power. While lensing power spectra are standard cosmological tools, non-Gaussianities induced by structure formation encode additional information in the bispectrum. In this work, we perform a comprehensive Fisher matrix forecast to quantify the detectability of these signals and the cosmological constraints achievable by combining power spectrum and bispectrum measurements from both CMB lensing and galaxy shear, including their cross-correlations. We model Stage-IV-like experimental specifications and forecast constraints on the standard $\Lambda$CDM parameters, the sum of neutrino masses ($\Sigma m_\nu$), and the dark energy equation of state ($w_0$). We demonstrate that incorporating bispectra significantly enhances constraints, reducing marginalized parameter uncertainties by up to $\sim \texttt{X}\%$ compared to power spectra alone. We project competitive constraints, including $\sigma(H_0) = \texttt{X.X~km/s/Mpc}$, $\sigma(\sigma_8) = \texttt{X.XXX}$, and $\sigma(\Sigma m_\nu) = \texttt{X.XX~eV}$.
    \end{abstract}
\pagebreak
%\tableofcontents

@@

Weak lensing has been used as a probe to constrain cosmological parameters since the 2000's \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies. They will do so much more accurately than previous surveys, enabling tighter constraints on parameters and progressing our understanding of physics as a whole. Typically, the statistic of interest is the power spectrum (or equivalently, two point function) of the lensing potential. This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime they all contain the same amount of information and can be trivially converted into one another, see appendix \ref{sec:weaklensstats}. In CMB surveys the temperature and polarization anisotropies are used to estimate the lensing potential \cite{Planck2018Lensing}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum. This has already been done for signals from galaxies \cite{vanWaerbeke2002}, but we do not yet have any proper detection of the CMB lensing bispectrum. Upcoming surveys such as the ones mentioned earlier may detect this signal. Therefore it is timely to consider the detectability, as well as ability to constrain cosmological parameter of CMB lensing power- and bispectra, galaxy lensing power- and bispectra, and cross correlations between the two weak lensing observables.

@@

Usefull derivations relevant to this paper can be found in ose, this paper also includes a derivation for the Fisher matrix formalism in the non-trivial context of bispectra with multiple tracers. The final formula can already be found in the literature (see e.g.\ the appendix of \cite{Kalaja_2021}), but the authors of this paper were not able to find a derivation and hence included it.

@@

\section{Theory}
\label{sec:theory}
\subsection{Weak lensing spectra}
Radiation from cosmological objects is distorted due to gravitational lensing. In the vast majority of cases this lensing is a relatively weak effect and is thus referred to as weak lensing. Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point on the sky and the true (unlensed) angle. This field is the gradient of the lensing potential $\psi(\hat {\mathbf n})$ given as a weighted integral over the mass distribution along the line of sight between earth and the observed point.

In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing to the expected unlensed signal, see e.g. \cite{cmblensingestimator}. In galaxy surveys, lensing alters the ellipiticities of observed galaxies. If a large enough number of galaxies are observed, this effect can be seperated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.

To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$. These are directly related to the matter power and bispectra as
\begin{empheq}[box=\fbox]{align*}
    C^{\psi_X\psi_Y}_l
    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(\ell/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
\end{empheq}
with $X, Y, Z \in \{\CMB, \gal\}$. A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}. $\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $C^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
$$
\begin{pmatrix}
    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
\end{pmatrix},
$$ 
the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
\begin{equation*}
    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
\end{equation*}
with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.

\subsection{Nonlinear matter bispectrum}
Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB. The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}. It is given by
\begin{equation}
    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm},
\end{equation}
where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
They are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln C^\delta_{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 C^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

\subsection{Fisher matrix analysis}
SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard. A full derivation of below statements can be found in appendix \ref{sec:fisher}. The auto and cross power spectrum Fisher matrix is given as
\begin{equation*}
    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l,
\end{equation*}
where
\begin{equation*}
    C_{l} := \begin{pmatrix}
        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
    \end{pmatrix}.
\end{equation*}
For the bispectra we instead get 
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
    \sum_{XYZ}\sum_{X'Y'Z'} 
    \partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
    (C^{-1})^{X X'}_{l_1}
    (C^{-1})^{Y Y'}_{l_2}
    (C^{-1})^{Z Z'}_{l_3}
    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
\end{equation*}
with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$. When only considering autospectra the formulas simplify to
\begin{equation*}
    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}\frac{2l+1}{2}\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
\end{equation*}
and (for bispectra)
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
    \frac{\partial_\alpha B^{X X X}_{l_1 l_2 l_3} 
    \partial_\beta B^{XXX}_{l_1 l_2 l_3}}{C^{XX}_{l_1}C^{XX}_{l_2}C^{XX}_{l_3}}.
\end{equation*}

\section{Calculation Details}\label{sec:calcdetails}
% \subsection{Numerical Considerations}
% % I used CAMB, how I took derivative, interpolation, fisher matrix sampling, matter bispectrum in terms of power spectrum

% Running cosmological simulations to obtain the matter power spectra and other functions was done with the CAMB package (REF).

% To speed up the calculation of the lensing powerspectra and especially the bispectra, an interpolation approach was used. Due to being 3 dimensional, it was not feasible to interpolate the lensing bispectrum directly. We instead chose to interpolate all 1 and 2 dimensional functions that the bispectra are based on and then calculate the bispectra with the interpolated functions instead. This significantly sped up the computation at the cost of small errors in the values in the bispectrum ($\sim 1\%$) when compared with the uninterpolated values.

% The time complexity of the Fisher matrix scales with the maximum multipole considered, $l_{max}$, as $\mathcal O (l_{max}^3)$, leading to time issues for realistic values of $l_{max}$. We thus calculated the $l_1 < l_2 < l_3$ part of each Fisher matrix element (see the appendix) by sampling every step of 10 %(CHECK IF THIS IS STILL CORRECT BEFORE PUBLISHING) 
% in $l_1$ and $l_2$. Due to the discontinuos nature of the bispectrum (e.g. it vanishes immediately if $l_1 + l_2 + l_3$ is and odd number) we didn't sample $l_3$ values. This was done similarly in (REF).

% \subsection{Choosing Experimental Parameters}

% lmin, lmax, galaxy redshift distribution, fiducial values (!), noise levels, 
% for SO, CMB - S4, Euclid, LSST
We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}, see table \ref{tab:fiducialpars}.
\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|}
        \hline
        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} \\
        \hline
        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ \\
        \hline
        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ \\
        \hline
        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$\\
        \hline
        Scalar spectral index & $n_s$ & 0.965 \\
        \hline
        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ \\
        \hline
        Sum of neutrino masses & $\sum m_\nu$ & $0.06$\\
        \hline
        Dark energy equation of state parameter & $w_0$ & -1 \\ \hline
    \end{tabular}
    \caption{Cosmological parameters considered ($\Lambda \text{CDM} + w_0 + m_\nu$) and their fiducial values used.}
    \label{tab:fiducialpars}
\end{table}

% For the parameter constraints we took the multipole range to be from 2 ($l_{min}$) to 2000 ($l_{max}$).

@@

\begin{table}[h]
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    \textbf{source} & \textbf{survey stage} & \textbf{noise vals} & \textbf{comparable experiments} \\
    \hline 
    \multirow{2}{*}{CMB} 
        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, SPT-3G \\
    \cline{2-4}
        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4, LiteBIRD \\
    \hline
    \multirow{2}{*}{galaxies} 
        & stage 3 & $\sigma_{\text{rms}} = 0.3$, $n_g = 5 \text{arcmin}^{-2}$ & DES, KiDS \\
    \cline{2-4}
        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, Euclid \\
    \hline
    \end{tabular}
    \caption{Noise levels considered for weak lensing of galaxies and the CMB. $\sigma$ (beam width) and $\Delta_P$ (polarization white noise) describe CMB survey specifications, while $\sigma_{\text{rms}}$ (intrinsic galaxy ellipticity) and $n_g$ (observed galaxy density) refer to galaxy shear surveys.}
    \label{tab:noiselevels}
\end{table}

>>>>

% !TEX root = ../paper/paper_draft.tex
\documentclass[11pt]{article} % Choose a document class, like 'article', 'report', or 'revtex4-2' for scientific papers

\usepackage{jcappub}

\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{multirow} % needed for merging cells vertically
\usepackage{booktabs}

@@

\usepackage[backend=bibtex]{biblatex}
\addbibresource{sources.bib}

@@

% Additional customizations (optional)
\setlength{\parindent}{0pt}   % No paragraph indentation
\setlength{\parskip}{1em}     % Add space between paragraphs
\title{Parameter Forecasts From CMB Lensing and Galaxy Lensing Power- and Bispectra}

\author[1]{Jonas Frugte}
\author[1]{P. Daniel Meerburg}

\affiliation[1]{Van Swinderen Institute for Particle Physics and Gravity, University of Groningen, Nijenborgh 4, 9747 AG Groningen, The Netherlands}
% \affil[ ]{}
% \affil[ ]{\textit{E-mail:} \texttt{jonasfrugte@gmail.com}}

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{}

\begin{document}

\maketitle

@@

%\tableofcontents

@@

Weak lensing has been used as a probe to constrain cosmological parameters since the 2000's \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aim to measure weak lensing of the CMB and of galaxies. These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of physics. Typically, the statistic of interest is the power spectrum (or equivalently, two point function) of the lensing potential. This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime they all contain the same amount of information and can be directly converted into one another, see appendix \ref{sec:weaklensstats}. In CMB surveys the temperature and polarization anisotropies are used to estimate the lensing potential \cite{Planck2018Lensing}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipiticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is thus to measure the lensing bispectrum. While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending. Upcoming surveys such as the ones mentioned earlier may detect this signal. Therefore it is timely to consider the detectability, as well as ability to constrain cosmological parameter of CMB lensing power- and bispectra, galaxy lensing power- and bispectra, and cross correlations between the two weak lensing observables.

@@

Useful derivations relevant to this paper can be found in the appendices. In particular, they include a derivation for the Fisher matrix formalism for the non-trivial context of bispectra with multiple tracers. While the final formula is available in the literature (e.g., the appendix of \cite{Kalaja_2021}), this paper includes a derivation as the authors were unable to locate one elsewhere.

@@

\section{Theory}
\label{sec:theory}
\subsection{Weak lensing spectra}
Radiation from cosmological objects is distorted due to gravitational lensing. Typically, this lensing is a relatively weak effect and is thus referred to as weak lensing. Weak lensing is quantified through the deflection field $\mathbf d(\hat {\mathbf n})$ which equals the difference between the observed angle of a point on the sky and the true (unlensed) angle. This field is the gradient of the lensing potential, $\psi(\hat {\mathbf n})$, which is expressed as a weighted integral of the mass distribution along the line of sight from the observer to the source.

In the case of CMB surveys, lensing alters the statistical properties of the temperature and polarization fields and can thus be calculated by comparing to the expected unlensed signal, see e.g. \cite{cmblensingestimator}. In galaxy surveys, lensing alters the ellipiticities of observed galaxies. If a large enough number of galaxies are observed, this effect can be separated from the intrinsic ellipiticities of the galaxies and thus the lensing potential can be estimated.

To constrain cosmological parameters we can then look at the lensing potential of the CMB, $\psi_\CMB$, and of galaxy surveys, $\psi_\gal$. These are directly related to the matter power and bispectra as
\begin{empheq}[box=\fbox]{align*}
    C^{\psi_X\psi_Y}_l
    & = \frac{9}{\ell^4}\Omega_m^2H_0^4\int_0^{\chi_*} \chi^2 \d\chi a(\eta_0-\chi)^{-2}W_X(\chi)W_Y(\chi)C^\delta(\ell/\chi,\eta_0-\chi),\\
    B^{\psi_X\psi_Y\psi_Z}_{\ell_1\ell_2\ell_3} &= \sqrt{\frac{(2l_1 + 1)(2l_2 + 1)(2l_3 + 1)}{4\pi}} \begin{pmatrix} l_1 & l_2 & l_3 \\ 0 & 0 & 0 \end{pmatrix} \frac{27}{l_1^2l_2^2l_3^2}\Omega_m^3H_0^6\\
    & \quad \times \int \chi^2\d \chi a(\eta_0-\chi)^{-3}W_X(\chi)W_Y(\chi)W_Z(\chi)  B^\delta(\{l_i/\chi\}, \eta_0-\chi),
\end{empheq}
with $X, Y, Z \in \{\CMB, \gal\}$. A derivation can be found in appendices \ref{sec:weaklensing} and \ref{sec:weaklensstats}, or in the literature, see e.g. \cite{bartelmann2001weak}. $\Omega_m$ is the present day matter density parameter, $H_0$ is the present day Hubble constant, $a(\eta)$ is the scale factor, $\eta_0$ is the conformal time today, $\chi$ is the comoving radial distance, $\chi_*$ the distance to surface of last scattering, $C^\delta(k, \eta)$ the matter power spectrum, $B^\delta(k_1, k_2, k_3, \eta)$ the matter bispectrum, $W_X(\chi)$ the window function, and
$$
\begin{pmatrix}
    l_1&l_2&l_3 \\ m_1 & m_2 & m_3
\end{pmatrix},
$$ 
the Wigner 3-j symbol. The only factor that differs between CMB and galaxy lensing is the window function, defined as
\begin{equation*}
    W_X(\chi) = \int_\chi^{\infty}\d\chi' p_X(\chi')\frac{\chi'-\chi}{\chi'\chi},
\end{equation*}
with $p_X$ the radius distribution of the radiation source. For the CMB we take $p(\chi) = \delta(\chi - \chi_*)$.

\subsection{Nonlinear matter bispectrum}
Determining the nonlinear matter powerspectrum can be done using numerical codes such as CAMB \cite{Lewis2000}. The nonlinear matter bispectrum was calculated from the powerspectrum using a fitting formula based on perturbation theory in \cite{bispfit}. It is given by
\begin{equation}
    B^\delta(k_1, k_2, k_3, \chi) = 2 F_2(k_1, k_2, z) C^\delta(k_1, z) C^\delta(k_2, z) + \text{2 perm},
\end{equation}
where \( C^\delta \) is the nonlinear matter power spectrum\footnote{Compare to the tree level bispectrum where we instead use the linear powerspectrum.}, and the kernel \( F_2 \) is modified from the tree level result with factors \( a(k, z) \), \( b(k, z) \), and \( c(k, z) \):
\begin{equation}
    F_2(k_1, k_2, z) = \frac{5}{7} a(k_1, z) a(k_2, z) + \frac{k_1^2 + k_2^2}{2 k_1 k_2} b(k_1, z) b(k_2, z) \cos \theta + \frac{2}{7} c(k_1, z) c(k_2, z) \cos^2 \theta.
\end{equation}
They are defined as:
\begin{equation}
    a(k, z) = \frac{1 + \sigma_8^{a_6}(z) \sqrt{0.7} Q(n_{\text{eff}}) (q^{a_1})^{n_{\text{eff}} + a_2}}{1 + (q^{a_1})^{n_{\text{eff}} + a_2}},
\end{equation}
\begin{equation}
    b(k, z) = \frac{1 + 0.2 a_3 (n_{\text{eff}} + 3) (q^{a_7})^{n_{\text{eff}} + 3 + a_8}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_8}},
\end{equation}
\begin{equation}
    c(k, z) = \frac{1 + \left[ \frac{4.5 a_4}{1.5 + (n_{\text{eff}} + 3)^4} \right] (q^{a_5})^{n_{\text{eff}} + 3 + a_9}}{1 + (q^{a_5})^{n_{\text{eff}} + 3.5 + a_9}}.
\end{equation}
Here, \( Q(n_{\text{eff}}) \) is given by:
\begin{equation}
    Q(x) = \frac{4 - 2x}{1 + 2^{x+1}}.
\end{equation}
The effective spectral index of the linear power spectrum is defined as:
\begin{equation}
    n_{\text{eff}} \equiv \frac{d \ln C^\delta_{\text{lin}}(k)}{d \ln k}.
\end{equation}
Additionally, \( q \) is given by:
\begin{equation}
    q = \frac{k}{k_{\text{NL}}},
\end{equation}
where \( k_{\text{NL}} \) is the scale at which nonlinearities become significant, satisfying:
\begin{equation}
    4 \pi k_{\text{NL}}^3 C^\delta_{\text{lin}}(k_{\text{NL}}, 0) = 1.
\end{equation}
The coefficients \( a_i \) are:
\begin{align*}
    a_1 &= 0.484, \quad a_2 = 3.740, \quad a_3 = -0.849, \quad a_4 = 0.392, \\
    a_5 &= 1.013, \quad a_6 = -0.575, \quad a_7 = 0.128, \quad a_8 = -0.722, \quad a_9 = -0.926.
\end{align*}

\subsection{Fisher matrix analysis}
SNR and parameter constraints are obtained through a Fisher matrix analysis as is standard \cite{dodelson2020modern}. A full derivation of below statements can be found in appendix \ref{sec:fisher}. The auto and cross power spectrum Fisher matrix is given as
\begin{equation*}
    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}(2l+1)\partial_\alpha C_l^{XY} (C^{-1})_l^{XX'}(C^{-1})_l^{YY'} \partial_\beta C^{X'Y'}_l,
\end{equation*}
where
\begin{equation*}
    C_{l} := \begin{pmatrix}
        C^{\psi_\kappa\psi_\kappa}_l & C^{\psi_\kappa\psi_\gamma}_l \\
        C^{\psi_\kappa\psi_\gamma}_l & C^{\psi_\gamma\psi_\gamma}_l
    \end{pmatrix}.
\end{equation*}
For the bispectra we instead get 
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
    \sum_{XYZ}\sum_{X'Y'Z'} 
    \partial_\alpha B^{X Y Z}_{l_1 l_2 l_3} 
    (C^{-1})^{X X'}_{l_1}
    (C^{-1})^{Y Y'}_{l_2}
    (C^{-1})^{Z Z'}_{l_3}
    \partial_\beta B^{X' Y' Z'}_{l_1 l_2 l_3},
\end{equation*}
with $\mathcal P_{l_1l_2l_3}$ defined as the number of distinct permutations that can be made with $l_1l_2l_3$. When only considering autospectra the formulas simplify to
\begin{equation*}
    F_{\alpha\beta} = \sum_{l} \sum_{XY}\sum_{X'Y'}\frac{2l+1}{2}\frac{\partial_{\alpha}C_l^{XX} \partial_\beta C^{XX}_l}{(C_l^{XX})^2},
\end{equation*}
and (for bispectra)
\begin{equation*}
    F_{\alpha\beta} = \sum_{l_1 \leq l_2 \leq l_3} \frac{\mathcal P _{l_1l_2l_3}}{6}
    \frac{\partial_\alpha B^{X X X}_{l_1 l_2 l_3} 
    \partial_\beta B^{XXX}_{l_1 l_2 l_3}}{C^{XX}_{l_1}C^{XX}_{l_2}C^{XX}_{l_3}}.
\end{equation*}

\section{Calculation Details}\label{sec:calcdetails}
% TODO: find proper references for stage 3 and stage 4 names as well as experimental parameters and real life surveys
We chose our fiducial cosmology in line with the results from the Planck mission \cite{planckresults}, see table \ref{tab:fiducialpars}. For all parameter constraints we use a prior similar to the one used in \cite{Planck2018Lensing}, see table \ref{tab:fiducialpars}. This prior only restricts $\Omega_b h^2$ and $n_s$ significantly. The power spectra, in particular, exhibit complete insensitivity to certain parameters. Consequently, relying solely on them to constrain the entire cosmological model results in very poor constraints. Future lensing surveys will of course use information from surveys measuring other statistics such as galaxy distributions or primary CMB anisotropies, so not using any priors is not representative of what stage 4 experiments could achieve. Even a weak prior improves the constraining power of our power spectra considerably.
\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        \textbf{parameter} & \textbf{notation} & \textbf{fiducial value} & \textbf{$\sigma$ of prior}\\
        \hline
        Hubble constant & $H_0$ & $67.4 \, \text{km/s/Mpc}$ & 17.3\\
        \hline
        Physical baryon density parameter & $\Omega_b h^2$ & $0.0223$ & 0.0005\\
        \hline
        Physical cold dark matter density parameter & $\Omega_c h^2$ & $0.119$ & $0.288$\\
        \hline
        Scalar spectral index & $n_s$ & 0.965 & 0.02 \\
        \hline
        Amplitude of primordial scalar fluctuations & $A_s$ & $2.13 \times 10^{-9}$ & $1\times 10^{-9}$ \\
        \hline
        Sum of neutrino masses & $\sum m_\nu$ & $0.06$ & $\infty$\\
        \hline
        Dark energy equation of state parameter & $w_0$ & -1 & $\infty$\\ \hline
    \end{tabular}
    \caption{Cosmological parameters varied ($\Lambda \text{CDM} + w_0 + m_\nu$). We use a prior similar to the one used by the Planck weak lensing results \cite{Planck2018Lensing} which essentially only restricts $\Omega_bh^2$ and $n_s$ significantly.}
    \label{tab:fiducialpars}
\end{table}

@@

\begin{table}[h!]
    \centering
    \begin{tabular}{|c|c|c|c|}
    \hline
    \textbf{source} & \textbf{survey stage} & \textbf{noise vals} & \textbf{comparable experiments} \\
    \hline 
    \multirow{2}{*}{CMB} 
        & stage 3 & $\sigma = 1'$, $\Delta_P = 6' \; \mu \text{K}$ & Advanced ACTPol, SPT-3G \\
    \cline{2-4}
        & stage 4 & $\sigma = 3'$, $\Delta_P = 1' \; \mu \text{K}$ & CMB-S4, LiteBIRD \\
    \hline
    \multirow{2}{*}{galaxies} 
        & stage 3 & $\sigma_{\text{rms}} = 0.3$, $n_g = 5 \text{arcmin}^{-2}$ & DES, KiDS \\
    \cline{2-4}
        & stage 4 & $\sigma_{\text{rms}} = 0.3$, $n_g = 30 \text{arcmin}^{-2}$ & LSST, Euclid \\
    \hline
    \end{tabular}
    \caption{Noise levels considered for weak lensing of galaxies and the CMB. $\sigma$ (beam width) and $\Delta_P$ (polarization white noise) describe CMB survey specifications, while $\sigma_{\text{rms}}$ (intrinsic galaxy ellipticity) and $n_g$ (observed galaxy density) refer to galaxy shear surveys.}
    \label{tab:noiselevels}
\end{table}
